Gradual Soundness: Lessons from Static Python

### 2021-10-27 ecoop-2022/paper.tex
ecoop: refine Array prose

For performance-critical applications, \SP{} includes a \emph{distinct} set of primitive types.
The basic primitives describe booleans and sized numbers, e.g., \code{int64}, \code{uint64}, \code{double}.
Two additional primitives, \code{Array} and \code{Vector}, describe collections.
Cinder implements these types with unboxed C numbers and C arrays, which cost
(and do!) much less than their Python counterparts.

@@

Primitives are carefully limited by the type system.
Neither a module-level nor a closure-level variable may have a primitive
type---because untyped code can mutate such variables.
Primitive types are incompatible with the dynamic type,
and there are no implicit upcasts from primitives
to a matching Python type.
Even a conjunction disallows mixing; for example, the \code{and} operator
requires either two Python booleans or two primitive booleans.

The only way for Python values and primitive values to interact
is through dedicated conversion functions.
Cinder handles conversions at the boundaries between \SP{} code
and untyped code to avoid a cascade of modify-then-run refactorings, but nowhere else.
Within typed code, programmers must satisfy the type checker with
appropriate conversions.

@@

\citet{pmmwplck-oopsla-2013} core calculus for Python language, foundational.

>>>>

For performance-critical applications, \SP{} includes a \emph{distinct} set of primitive types
that describe booleans and sized numbers, e.g., \code{int64}, \code{uint64}, \code{double}.
Cinder implements these types with unboxed C values, which are much simpler and cheaper
than their Python counterparts.
\SP{} also includes two special datatypes, \code{Array} and \code{Vector}, that store primitives efficiently.
These are not themselves primitive.

Primitives are carefully limited by the type system.
Neither a module-level nor a closure-level variable may have a primitive
type---because untyped code can mutate such variables.
Primitive types are incompatible with the dynamic type,
and there are no implicit upcasts from primitives
to a matching Python type.
Even a conjunction disallows mixing; for example, the \code{and} operator
requires either two Python booleans or two primitive booleans.
The only way for Python values and primitive values to interact
is through dedicated conversion functions.
Cinder handles conversions at the boundaries between \SP{} code
and untyped code to avoid a cascade of modify-then-run refactorings, but nowhere else.
Within typed code, programmers must satisfy the type checker with
appropriate conversions.

@@

@@

Full Monty(?)~\cite{pmmwplck-oopsla-2013} core calculus for Python language, foundational.
